---
title: "Animacy agreement in non-verbal predication in Botlikh"
subtitle: "average scores, linear models and graphs"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)

#packages

#data manipulation
library(tidyverse)

#linear models
library(lme4)
library(lmerTest)

#plots
library(effects)
library(corrplot)
library(wesanderson)

```

```{r}
#data
setwd("/home/samira/Git/2021mt_negcop")
dat <- read_tsv("data.csv")

no_prof <- dat[-1,]

dat %>% pivot_longer(cols = starts_with("speaker"), values_to = "score") -> dat_long

proficiency <- (dat %>% filter(question == "proficiency") %>% t() %>% as.data.frame() %>% drop_na())[-1,] %>% as.data.frame() %>% rownames_to_column(var = "speaker") %>% filter(speaker != "total")
proficiency <- data.frame(speaker=proficiency$speaker, proficiency = proficiency$`.`)

```

# Preliminary

## Speaker variation

```{r}

corrplot(cor(dat %>% select(starts_with("speaker"))), method = 'color')

```

## Score deviation by speaker proficiency (yes/no)

```{r}

t.test(score_sd ~ proficiency, dat_long %>% group_by(name) %>% summarise(score_sd = sd(score)) %>% left_join(proficiency, by=c('name'= 'speaker')))

```

## Average score of correct examples

```{r}
model <- lmer(score ~ correct + (1 + correct|name), data=dat_long, control = lmerControl(optimizer="bobyqa"))
summary(model)
```

```{r}
plot(predictorEffect("correct", model))
```

# Score of each auxiliary with animate/inanimate referents

```{r}

aux_nr <- no_prof %>%
  select(aux, animate)%>%
  group_by(aux, animate) %>%
  summarise(nr = n())

aux_score <- no_prof %>%
  select(aux, animate, avg)%>%
  group_by(aux, animate) %>%
  summarise(score = sum(avg))

aux_score$nr <- aux_nr$nr
aux_score$avg_score <- (aux_score$score / aux_score$nr)

aux_score

```

# Innovative system

## Inanimate auxiliary

```{r}

books <- dat %>%
  filter(subject_en %in% c("book", "books")) %>%
  filter(system == "innovative") %>%
  select(subject_en, aux, avg)

books

```

## Animate auxiliary

```{r}

li <- no_prof[(no_prof$aux == "лъич1и"),]

li_nr <- li %>%
  select(subject_en, number, avg) %>%
  group_by(subject_en, number) %>%
  summarise(nr=n())

li_score <- li %>%
  select(subject_en, number, avg) %>%
  group_by(subject_en, number) %>%
  summarise(score=sum(avg))

li_score$nr <- li_nr$nr
li_score$avg_score <- (li_score$score / li_score$nr)

li_score

```

### Number effect

```{r}

model <- lm(score ~ number, 
              data=dat_long %>% filter(aux_value == "an", animate=="yes"))
summary(model)

```

```{r}
plot(Effect("number", model))
```

### Animate category effect

```{r}

model <- lm(score ~ human, 
              data=dat_long %>% filter(aux_value == "an", animate=="yes"))
summary(model)

```

```{r}
plot(Effect("human", model))
```

## Distribution of innovative auxiliaries

```{r}

innov <- no_prof[(no_prof$system == "innovative"),]

q2 <- innov %>%
  mutate(subject = fct_reorder(subject_en, desc(avg))) %>%
  ggplot(aes(x=subject, y=avg))+
  geom_line(aes(colour = aux, group = aux))+
  geom_point(aes(colour = aux))+
  scale_color_manual(values=wes_palette(n=3, name="Moonrise3"))+
  theme_classic()

q2 + theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust=1),
           axis.title.x=element_blank(), 
           axis.title.y=element_blank())

```

# Inherited

## Compared to innovative

```{r}

an <- no_prof %>%
  filter(aux %in% c("лъич1и", "рук1ич1а"))

q3 <- ggplot(an, aes(fct_reorder(subject_en, desc(avg)), avg)) + 
  geom_line(aes(colour = aux, group = aux))+
  geom_point(aes(colour = aux))+
  scale_color_manual(values=c("#006B3C", "#8A2BE2"))+
  theme_classic()

q3 + theme(axis.text.x = element_text(angle = 50, vjust = 1, hjust=1),
           axis.title.x=element_blank(), 
           axis.title.y=element_blank())

```

```{r}

model <- lm(score ~ system, 
              data=dat_long %>% filter(animate=="yes"))
summary(model)

```

## Animate category

### For the animate plural

```{r}
model <- lm(score ~ human, 
              data=dat_long %>% filter(aux_value == "anpl"))
summary(model)
```

### For the neuter singular / inanimate plural form

```{r}
model <- lm(score ~ human, 
              data=dat_long %>% filter(aux_value == "n"))
summary(model)
```

## Number

compare avg_score for correct = yes + number = pl

```{r}

inherited_nr <- no_prof %>%
  filter(system == "inherited") %>%
  select(aux, correct, number, avg) %>%
  group_by(aux, correct, number) %>%
  summarise(nr=n())

inherited_nrscore <- no_prof %>%
  filter(system == "inherited") %>%
  select(aux, correct, number, avg) %>%
  group_by(aux, correct, number) %>%
  summarise(score=sum(avg))

inherited_nrscore$nr <- inherited_nr$nr
inherited_nrscore$avg_score <- (inherited_nrscore$score / inherited_nrscore$nr)

inherited_nrscore

```



